```{r}
library(dplyr)
library(ggplot2)
library(splines)
library(data.table)
library(caret)
library('manipulate')
```

t = cancer
n = healthy


### א.

```{r}
path = "/Users/lisabensoussan/Desktop/FinalLab"
#path = "/Users/Emmanuelle Fareau/Downloads"

load(sprintf("%s/reads_100_A1.rda", path))
load(sprintf("%s/reads_100_B1.rda", path))
load(sprintf("%s/GC_100.rda", path))

# Assuming each index covers 100K and 250-300 index covers 25M-30M
index_range <- 0:700000  # Adjust this based on your actual data indexing

# Filter data for the 25M-30M region
filtered_reads_cancer <- reads_100_A1[index_range]
filtered_reads_healthy <- reads_100_B1[index_range]
filtered_gc <- GC_100[index_range]

# Reshape data for plotting
chr1_can = matrix(filtered_reads_cancer, nr = 50)
reads_5K_c = colSums(chr1_can)

chr1_hea = matrix(filtered_reads_healthy, nr = 50)
reads_5K_h = colSums(chr1_hea)

gc_mat = matrix(filtered_gc, nr = 50)
GC_5K = colMeans(gc_mat)

data_sample <- data.frame(GC = GC_5K, Reads_Healthy = reads_5K_h, Reads_Cancer = reads_5K_c)

```

```{r}
library(ggplot2)
library(gridExtra)  # for arranging plots

# Plot for Healthy Samples
p1 <- ggplot(data_sample, aes(x = GC, y = Reads_Healthy)) +
  geom_point(color = 'lightblue', size = 0.5) +
  labs(title = "GC Content vs. Reads for Healthy Samples", x = "GC Content", y = "Reads") +
  theme_minimal() +
  scale_x_continuous(limits = c(0.2, 0.75)) +
  scale_y_continuous(limits = c(0, 600))

# Plot for Cancer Samples
p2 <- ggplot(data_sample, aes(x = GC, y = Reads_Cancer)) +
  geom_point(color = 'lightpink', size = 0.5) +
  labs(title = "GC Content vs. Reads for Cancer Samples", x = "GC Content", y = "Reads") +
  theme_minimal() +
  scale_x_continuous(limits = c(0.2, 0.75)) +
  scale_y_continuous(limits = c(0, 600))

# Arrange plots vertically for comparison
grid.arrange(p1, p2, ncol = 1)

```

```{r}
library(splines)

l_1h = length(reads_5K_h)
l_1c = length(reads_5K_c)

GC=GC_5K[1:l_1h]
mod_1h = lm(reads_5K_h~bs(GC,deg = 3,knots = c(seq(0.25,0.6,0.025),0.65)),subset = (reads_5K_h>50 & reads_5K_h<600 ))


pred_range = seq(0.2,0.75,0.01)
resp_h = predict(mod_1h,list(GC=pred_range))

preds1h = pmax(predict(mod_1h,list(GC=GC_5K)),0)




GC=GC_5K[1:l_1c]
mod_1c = lm(reads_5K_c~bs(GC,deg = 3,knots = c(seq(0.25,0.6,0.025),0.65)),subset = (reads_5K_c>50) & (reads_5K_c<550))

pred_range = seq(0.2,0.75,0.01)
resp_c = predict(mod_1c,list(GC=pred_range))
preds1c = pmax(predict(mod_1c,list(GC=GC_5K)),0)

```
```{r}
```






```{r}
library(ggplot2)


# Define the correction function
one_sample_correct <- function(samp, preds, eps) {
  one_correct = (samp + eps) / (preds + eps)
  adj_med = 1 / median(one_correct)
  return(adj_med * one_correct)
}

# Define epsilon
eps = 1

# start_index = 250*3  #Bins of 5000/3=1666
# end_index = 300*3

# Applying the correction and uncorrection to the specific segment
uncorrected_1h = one_sample_correct(reads_5K_h, rep(1, length(reads_5K_h)), eps)

uncorrected_1c = one_sample_correct(reads_5K_c, rep(1, length(reads_5K_c)), eps)

data_healthy <- data.frame(
  SegmentIndex = seq( length.out = length(uncorrected_1h)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1h
)

data_cancer <- data.frame(
  SegmentIndex = seq(length.out = length(uncorrected_1c)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1c
)


# Plotting Healthy Samples
p1 <- ggplot(data_healthy, aes(x = SegmentIndex, y = UncorrectedReads)) +
  geom_point(color = "lightblue", size=0.5) +
  labs(title = "Healthy Samples: Uncorrected Read Distribution",
       x = "Segment Index", y = "Read Count") +
  theme_minimal() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black")+  # Reference line at y = 1
  ylim(c(0,3))

# Plotting Cancer Samples
p2 <- ggplot(data_cancer, aes(x = SegmentIndex, y = UncorrectedReads)) +
  geom_point(color = "lightpink", size=0.5) +
  labs(title = "Cancer Samples: Uncorrected Read Distribution",
       x = "Segment Index", y = "Read Count") +
  theme_minimal() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black") + # Reference line at y = 1
  ylim(c(0,3))

# Using gridExtra to arrange the plots
library(gridExtra)
grid.arrange(p1, p2, ncol = 1)

```

### ב. 


```{r}
knots <- quantile(GC_5K, probs = c(0.25, 0.5, 0.75))

model_h <- lm(reads_5K_h ~ ns(GC_5K, knots = knots), data = data_healthy)
model_c <- lm(reads_5K_c ~ ns(GC_5K, knots = knots), data = data_cancer)


estimate_copy_number_h <- function(Y, GC, f) {
  # Predict using the model and create new data frame with the name used in the model
  f_gc_hat <- predict(model_h, newdata = data.frame(GC_5K = GC))
  # Apply the random noise multiplier
  noise_factor <- rnorm(length(Y), 1, 0.1)
  a_hat <- Y / (f_gc_hat * noise_factor)
  return(a_hat)
}

estimate_copy_number_c <- function(Y, GC, f) {
  # Predict using the model and create new data frame with the name used in the model
  f_gc_hat <- predict(model_c, newdata = data.frame(GC_5K = GC))
  # Apply the random noise multiplier
  noise_factor <- rnorm(length(Y), 1, 0.1)
  a_hat <- Y / (f_gc_hat * noise_factor)
  return(a_hat)
}

Y_h <- reads_5K_h
Y_c <- reads_5K_c


estimated_copy_numbers_h <- estimate_copy_number_h(Y_h, GC_5K, model_h)
estimated_copy_numbers_c <- estimate_copy_number_c(Y_c, GC_5K, model_c)

head(estimated_copy_numbers_h)
head(estimated_copy_numbers_c)

```

```{r}
library(ggplot2)

# Data frame for healthy samples
data_healthy_plot <- data.frame(
  Reads = reads_5K_h,
  EstimatedCopies = estimated_copy_numbers_h
)

# Data frame for cancer samples
data_cancer_plot <- data.frame(
  Reads = reads_5K_c,
  EstimatedCopies = estimated_copy_numbers_c
)

# Plotting for Healthy Samples
p1 <- ggplot(data_healthy_plot, aes(x = Reads, y = EstimatedCopies)) +
  geom_point(color = 'lightblue', alpha = 0.6, size=0.5) +  # Points with some transparency
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 3), se = FALSE, color = "black", size=0.5) +  # Spline regression
  labs(title = "Healthy Samples: Reads vs. Estimated Copy Numbers",
       x = "Reads",
       y = "Estimated Copy Numbers") +
  theme_minimal()+
  ylim(c(-1, 4))

# Plotting for Cancer Samples
p2 <- ggplot(data_cancer_plot, aes(x = Reads, y = EstimatedCopies)) +
  geom_point(color = 'lightpink', alpha = 0.6, size=0.5) +  # Points with some transparency
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 3), se = FALSE, color = "black", size=0.5) +  # Spline regression
  labs(title = "Cancer Samples: Reads vs. Estimated Copy Numbers",
       x = "Reads",
       y = "Estimated Copy Numbers") +
  theme_minimal()+
  ylim(c(-1, 4))

# Using gridExtra to arrange the plots if needed
library(gridExtra)
grid.arrange(p1, p2, ncol = 1)

```

### ג.

```{r}
library(ggplot2)

# Create a data frame for plotting
data_plot <- data.frame(
  GC = GC_5K,
  Reads_Healthy = reads_5K_h,
  Reads_Cancer = reads_5K_c
)

# Plotting GC content effect using spline regression
ggplot(data_plot, aes(x = GC, y = Reads_Healthy)) +
  geom_point(color = 'lightblue', alpha = 0.5) +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, df = 3), se = TRUE, color = "black") +
  labs(title = "Modeling GC Count Effect on Reads (Healthy Samples)",
       x = "GC Count",
       y = "Reads") +
  theme_minimal()

# Repeat for cancer samples if needed
ggplot(data_plot, aes(x = GC, y = Reads_Cancer)) +
  geom_point(color = 'lightpink', alpha = 0.5) +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, df = 3), se = TRUE, color = "black") +
  labs(title = "Modeling GC Count Effect on Reads (Cancer Samples)",
       x = "GC Count",
       y = "Reads") +
  theme_minimal()

```

```{r}

# start_index = 250*3  #Bins of 5000/3=1666
# end_index = 300*3

# Applying the correction and uncorrection to the specific segment
uncorrected_1h = one_sample_correct(reads_5K_h, rep(1, length(reads_5K_h)), eps)

uncorrected_1c = one_sample_correct(reads_5K_c, rep(1, length(reads_5K_c)), eps)

data_healthy <- data.frame(
  SegmentIndex = seq(25, 30, length.out = length(uncorrected_1h)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1h
)

data_cancer <- data.frame(
  SegmentIndex = seq(25, 30, length.out = length(uncorrected_1c)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1c
)
```

### ה. 

```{r}
# Function to calculate RMSE based on the given formula
calculate_rmse <- function(estimated_copies) {
  n <- length(estimated_copies)
  rmse <- sqrt(sum((estimated_copies - 1)^2) / n)
  return(rmse)
}

# Assuming 'estimated_copy_numbers_h' and 'estimated_copy_numbers_c' are your estimated copy numbers for healthy and cancer samples
rmse_healthy <- calculate_rmse(estimated_copy_numbers_h)
rmse_cancer <- calculate_rmse(estimated_copy_numbers_c)

# Output the RMSE values
print(paste("RMSE for Healthy Samples:", rmse_healthy))
print(paste("RMSE for Cancer Samples:", rmse_cancer))

```


```{r}
# Assuming 'estimated_copy_numbers_h' and 'estimated_copy_numbers_c' contain the estimated copy numbers
# for healthy and cancer samples respectively, and true values Y_i are all 1

# Calculate the absolute errors
absolute_errors_healthy = abs(1 - estimated_copy_numbers_h)
absolute_errors_cancer = abs(1 - estimated_copy_numbers_c)

# Calculate the Median Absolute Error (MAE)
mae_healthy = median(absolute_errors_healthy)
mae_cancer = median(absolute_errors_cancer)

# Print the MAE for both samples
print(paste("Median Absolute Error for Healthy Samples:", mae_healthy))
print(paste("Median Absolute Error for Cancer Samples:", mae_cancer))

```


### ו.

```{r}
two_sample_correct_e = function(cancer, healthy,eps) {
  two_correct = (cancer+eps)/(healthy+eps)
  two_correct_aft_med = two_correct/median(two_correct,na.rm=TRUE)
  return(two_correct_aft_med)
}

two_sample_wout_gc = two_sample_correct_e(uncorrected_1c,uncorrected_1h,0.01)
library(ggplot2)
library(scales)  # for the label formatting

# Assuming two_sample_wout_gc is already computed as shown earlier
# Create a data frame with an index for plotting
data_plot <- data.frame(
  SegmentIndex = 1:length(two_sample_wout_gc),
  CorrectedRatio = two_sample_wout_gc
)

# Convert index to genomic position assuming the index spans 25M to 30M
# Let's say the total range covers 5M bases and you have N points
total_points <- length(two_sample_wout_gc)
base_per_point <- 5000000 / total_points  # 5M bases total range

data_plot$GenomicPosition <- (data_plot$SegmentIndex - 1) * base_per_point  # 25M start

# Plot using ggplot
p1 <- ggplot(data_plot, aes(x = GenomicPosition, y = CorrectedRatio)) +
  geom_point(color = 'darkorchid1', alpha = 0.5, size=0.5) +  # Smaller points; adjust size as needed
  geom_hline(yintercept = 1, color = "black", linetype = "dashed") +  # Reference line at y = 1
  labs(title = "Two Sample Correction without GC", x = "Genomic Position", y = "Read Count") +
  theme_minimal() +
  ylim(0, 2) +  # Set limits for y-axis
  scale_x_continuous(labels = comma)  # Formatting x-axis labels with commas for readability



# Combining the datasets for plotting
combined_data <- rbind(data_healthy, data_cancer)
combined_data$Type <- rep(c("Healthy", "Cancer"), each = nrow(data_healthy))

# Creating the plot
combined_plot <- ggplot(combined_data, aes(x = SegmentIndex, y = UncorrectedReads, color = Type)) +
  geom_point(alpha = 0.3, size=0.5) +  # Plot points
  scale_color_manual(values = c("Healthy" = "lightblue", "Cancer" = "lightpink")) +  # Set colors
  labs(title = "One sample Correction Read Distribution",
       x = "Segment Index", y = "Read Count") +
  theme_minimal() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black") +  # Reference line at y = 1
  ylim(c(0, 3))  # Set y-axis limits



# Print the plot
print(p1)
print(combined_plot)

grid.arrange(p1, combined_plot, ncol = 1)


```
POURQUOI AUTOURS DE 1 ?

In the plot you're referring to, all the values cluster around the value 1 due to the method of data normalization used, specifically designed to standardize the read count distribution across different samples.

### Explanation of Normalization Around 1:

1. **Purpose of Normalization:**
   Normalization is commonly used in genomic data analysis to adjust for various biases and scaling issues between samples. The goal is to make the datasets comparable by aligning them on a common scale.

2. **Median Adjustment:**
   The method likely applied here involves adjusting each observation by a factor that normalizes the median of the adjusted values to 1. This type of normalization ensures that the central tendency (median) of the read counts across different segments is consistent, helping to compare across different conditions like healthy vs. cancer samples without the interference of scale differences.

3. **Use of Epsilon in Normalization:**
   Epsilon smoothing might have been employed to stabilize the ratios, especially in segments where predicted values (e.g., based on a model or an external control) are extremely low or zero. By adding a small constant (epsilon) to both numerator and denominator, the ratio becomes more stable and less sensitive to small fluctuations in the data, reducing the impact of outliers and avoiding undefined expressions due to division by zero.

4. **Application of Normalization Across Segments:**
   Each point in the plot represents the normalized read count for a specific segment of the genome. By scaling these values around 1, it becomes easier to visually inspect the plot for anomalies or differences between healthy and cancer samples. Values significantly above or below 1 might indicate genomic regions with unusual activity or variation, such as amplifications or deletions that could be relevant to cancer research.

5. **Visual Inspection:**
   The dashed line at y=1 acts as a reference, making it straightforward to see which data points deviate from this expected baseline. This setup is particularly useful for quickly identifying segments where the read counts are unusually high or low, which could warrant further biological investigation.

In summary, this normalization approach provides a standardized way to compare read counts across different samples or conditions, highlighting deviations that might be biologically significant.\




### ה. 



```{r}
index_range <- 260000:280000  # Adjust this based on your actual data indexing

# Filter data for the 25M-30M region
filtered_reads_cancer <- reads_100_A1[index_range]
filtered_reads_healthy <- reads_100_B1[index_range]
filtered_gc <- GC_100[index_range]

# Reshape data for plotting
chr1_can = matrix(filtered_reads_cancer, nr = 50)
reads_5K_c = colSums(chr1_can)

chr1_hea = matrix(filtered_reads_healthy, nr = 50)
reads_5K_h = colSums(chr1_hea)

gc_mat = matrix(filtered_gc, nr = 50)
GC_5K = colMeans(gc_mat)

data_sample <- data.frame(GC = GC_5K, Reads_Healthy = reads_5K_h, Reads_Cancer = reads_5K_c)
```


```{r}



l_1h = length(reads_5K_h)
l_1c = length(reads_5K_c)

GC=GC_5K[1:l_1h]
mod_1h = lm(reads_5K_h~bs(GC,deg = 3,knots = c(seq(0.25,0.6,0.025),0.65)),subset = (reads_5K_h>50 & reads_5K_h<600 ))


pred_range = seq(0.2,0.75,0.01)
resp_h = predict(mod_1h,list(GC=pred_range))

preds1h = pmax(predict(mod_1h,list(GC=GC_5K)),0)




GC=GC_5K[1:l_1c]
mod_1c = lm(reads_5K_c~bs(GC,deg = 3,knots = c(seq(0.25,0.6,0.025),0.65)),subset = (reads_5K_c>50) & (reads_5K_c<550))

pred_range = seq(0.2,0.75,0.01)
resp_c = predict(mod_1c,list(GC=pred_range))
preds1c = pmax(predict(mod_1c,list(GC=GC_5K)),0)
```




```{r}

# Define the correction function
one_sample_correct <- function(samp, preds, eps) {
  one_correct = (samp + eps) / (preds + eps)
  adj_med = 1 / median(one_correct)
  return(adj_med * one_correct)
}

# Define epsilon
eps = 1

# start_index = 250*3  #Bins of 5000/3=1666
# end_index = 300*3

# Applying the correction and uncorrection to the specific segment
uncorrected_1h = one_sample_correct(reads_5K_h, rep(1, length(reads_5K_h)), eps)

uncorrected_1c = one_sample_correct(reads_5K_c, rep(1, length(reads_5K_c)), eps)

data_healthy <- data.frame(
  SegmentIndex = seq( length.out = length(uncorrected_1h)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1h
)

data_cancer <- data.frame(
  SegmentIndex = seq(length.out = length(uncorrected_1c)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1c
)

```



```{r}
knots <- quantile(GC_5K, probs = c(0.25, 0.5, 0.75))

model_h <- lm(reads_5K_h ~ ns(GC_5K, knots = knots), data = data_healthy)
model_c <- lm(reads_5K_c ~ ns(GC_5K, knots = knots), data = data_cancer)


estimate_copy_number_h <- function(Y, GC, f) {
  # Predict using the model and create new data frame with the name used in the model
  f_gc_hat <- predict(model_h, newdata = data.frame(GC_5K = GC))
  # Apply the random noise multiplier
  noise_factor <- rnorm(length(Y), 1, 0.1)
  a_hat <- Y / (f_gc_hat * noise_factor)
  return(a_hat)
}

estimate_copy_number_c <- function(Y, GC, f) {
  # Predict using the model and create new data frame with the name used in the model
  f_gc_hat <- predict(model_c, newdata = data.frame(GC_5K = GC))
  # Apply the random noise multiplier
  noise_factor <- rnorm(length(Y), 1, 0.1)
  a_hat <- Y / (f_gc_hat * noise_factor)
  return(a_hat)
}

Y_h <- reads_5K_h
Y_c <- reads_5K_c


estimated_copy_numbers_h <- estimate_copy_number_h(Y_h, GC_5K, model_h)
estimated_copy_numbers_c <- estimate_copy_number_c(Y_c, GC_5K, model_c)

head(estimated_copy_numbers_h)
head(estimated_copy_numbers_c)

```

```{r}


# Data frame for healthy samples
data_healthy_plot <- data.frame(
  Reads = reads_5K_h,
  EstimatedCopies = estimated_copy_numbers_h
)

# Data frame for cancer samples
data_cancer_plot <- data.frame(
  Reads = reads_5K_c,
  EstimatedCopies = estimated_copy_numbers_c
)


```


```{r}

# Create a data frame for plotting
data_plot <- data.frame(
  GC = GC_5K,
  Reads_Healthy = reads_5K_h,
  Reads_Cancer = reads_5K_c
)


```

```{r}

# start_index = 250*3  #Bins of 5000/3=1666
# end_index = 300*3

# Applying the correction and uncorrection to the specific segment
uncorrected_1h = one_sample_correct(reads_5K_h, rep(1, length(reads_5K_h)), eps)

uncorrected_1c = one_sample_correct(reads_5K_c, rep(1, length(reads_5K_c)), eps)

data_healthy <- data.frame(
  SegmentIndex = seq(25, 30, length.out = length(uncorrected_1h)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1h
)

data_cancer <- data.frame(
  SegmentIndex = seq(25, 30, length.out = length(uncorrected_1c)),  # Transform to 25M to 30M scale
  UncorrectedReads = uncorrected_1c
)
```


### ו.

```{r}
two_sample_correct_e = function(cancer, healthy,eps) {
  two_correct = (cancer+eps)/(healthy+eps)
  two_correct_aft_med = two_correct/median(two_correct,na.rm=TRUE)
  return(two_correct_aft_med)
}

two_sample_wout_gc = two_sample_correct_e(uncorrected_1c,uncorrected_1h,0.01)
library(ggplot2)
library(scales)  # for the label formatting

# Assuming two_sample_wout_gc is already computed as shown earlier
# Create a data frame with an index for plotting
data_plot <- data.frame(
  SegmentIndex = 1:length(two_sample_wout_gc),
  CorrectedRatio = two_sample_wout_gc
)

# Convert index to genomic position assuming the index spans 25M to 30M
# Let's say the total range covers 5M bases and you have N points
total_points <- length(two_sample_wout_gc)
base_per_point <- 5000000 / total_points  # 5M bases total range

data_plot$GenomicPosition <- (data_plot$SegmentIndex - 1) * base_per_point  # 25M start



# Combining the datasets for plotting
combined_data <- rbind(data_healthy, data_cancer)
combined_data$Type <- rep(c("Healthy", "Cancer"), each = nrow(data_healthy))


```




```{r}
# Function to calculate RMSE based on the given formula
calculate_rmse <- function(estimated_copies) {
  n <- length(estimated_copies)
  rmse <- sqrt(sum((estimated_copies - 1)^2) / n)
  return(rmse)
}

# Assuming 'estimated_copy_numbers_h' and 'estimated_copy_numbers_c' are your estimated copy numbers for healthy and cancer samples
rmse_healthy <- calculate_rmse(estimated_copy_numbers_h)
rmse_cancer <- calculate_rmse(estimated_copy_numbers_c)

# Output the RMSE values
print(paste("RMSE for Healthy Samples:", rmse_healthy))
print(paste("RMSE for Cancer Samples:", rmse_cancer))

```


```{r}
# Assuming 'estimated_copy_numbers_h' and 'estimated_copy_numbers_c' contain the estimated copy numbers
# for healthy and cancer samples respectively, and true values Y_i are all 1

# Calculate the absolute errors
absolute_errors_healthy = abs(1 - estimated_copy_numbers_h)
absolute_errors_cancer = abs(1 - estimated_copy_numbers_c)

# Calculate the Median Absolute Error (MAE)
mae_healthy = median(absolute_errors_healthy)
mae_cancer = median(absolute_errors_cancer)

# Print the MAE for both samples
print(paste("Median Absolute Error for Healthy Samples:", mae_healthy))
print(paste("Median Absolute Error for Cancer Samples:", mae_cancer))
```


```{r}


knots <- quantile(GC_5K, probs = c(0.25, 0.5, 0.75))

model_h <- lm(reads_5K_h ~ ns(GC_5K, knots = knots), data = uncorrected_1h)
model_c <- lm(reads_5K_c ~ ns(GC_5K, knots = knots), data = uncorrected_1c)


estimate_copy_number_h <- function(Y, GC, f) {
  # Predict using the model and create new data frame with the name used in the model
  f_gc_hat <- predict(model_h, newdata = data.frame(GC_5K = GC))
  # Apply the random noise multiplier
  noise_factor <- rnorm(length(Y), 1, 0.1)
  a_hat <- Y / (f_gc_hat * noise_factor)
  return(a_hat)
}

estimate_copy_number_c <- function(Y, GC, f) {
  # Predict using the model and create new data frame with the name used in the model
  f_gc_hat <- predict(model_c, newdata = data.frame(GC_5K = GC))
  # Apply the random noise multiplier
  noise_factor <- rnorm(length(Y), 1, 0.1)
  a_hat <- Y / (f_gc_hat * noise_factor)
  return(a_hat)
}

Y_h <- reads_5K_h
Y_c <- reads_5K_c


estimated_copy_numbers_h <- estimate_copy_number_h(Y_h, GC_5K, model_h)
estimated_copy_numbers_c <- estimate_copy_number_c(Y_c, GC_5K, model_c)

head(estimated_copy_numbers_h)
head(estimated_copy_numbers_c)

calculate_rmse <- function(estimated_copies) {
  n <- length(estimated_copies)
  rmse <- sqrt(sum((estimated_copies - 1)^2) / n)
  return(rmse)
}

# Assuming 'estimated_copy_numbers_h' and 'estimated_copy_numbers_c' are your estimated copy numbers for healthy and cancer samples
rmse_healthy <- calculate_rmse(estimated_copy_numbers_h)
rmse_cancer <- calculate_rmse(estimated_copy_numbers_c)

print(rmse_healthy)
print(rmse_cancer)

```
```{r}
model_corr <- lm(reads_5K_h ~ ns(GC_5K, knots = knots), data = data_plot)
estimate_copy_number_corr <- function(Y, GC, f) {
  # Predict using the model and create new data frame with the name used in the model
  f_gc_hat <- predict(model_corr, newdata = data.frame(GC_5K = GC))
  # Apply the random noise multiplier
  noise_factor <- rnorm(length(Y), 1, 0.1)
  a_hat <- Y / (f_gc_hat * noise_factor)
  return(a_hat)
}

estimated_copy_numbers_cor <- estimate_copy_number_corr(Y_c, GC_5K, model_corr)
rmse_corr <- calculate_rmse(estimated_copy_numbers_cor)
print(rmse_corr)

```

